Data Flow Analysis\\in the Presence of Correlated Calls

### 2014-12-09 content/intro.tex
comments

Data-flow analysis computes an approximation of how values may flow  
through a program, and has applications in compiler optimization, 
programming tools, and computer security, and many other areas.
Data-flow analyses operate on \textit{control-flow graphs} (CFGs) that model 
the order in which the instructions of a program are executed. This is
typically done by associating \textit{flow functions} that represent how 
data is propagated with the edges of the control-flow graph.  A 
\textit{confluence operator} specifies how the data facts that have been 
computed along different paths should be merged when the paths join.

@@

We present a formalization of the transformation and prove its correctness
as follows. First, we derive an ``Equivalence IDE problem'' from the original 
IFDS problem by associating the identity environment functions with all edges,
and show that the solution to this Equivalence IDE problem can be mapped back
to the solution of the original IFDS problem. Then, we derive a ``Correlated Calls 
IDE Problem'' from the original IFDS problem and show how a solution to this problem 
can be mapped to the solution to the original IFDS problem, but also how a
more precise IFDS result can be derived from it. 
\otodo{Do we need to mention the imprecise map from the CC-IDE result to the IFDS result? It
doesn't really seem relevant, and could be confusing.}
We also show that the correlated-calls analysis is sound, i.e., that it never considers concrete execution paths as infeasible.
This is illustrated schematically
in Figure~\ref{fig:transformations}%
\footnote{ 
  The labels $\result$, $\backEq$, $\backCC$, $\transEq$, and $\transCC$ on edges in 
  Figure~\ref{fig:transformations} reflect a number of mappings and projections 
  that will be defined in Section~\TODO{ref} and that can be ignored here.
}.  
\ftodo{would it make sense to add an edge in Figure~\ref{fig:transformations} from the IFDS problem to the IFDS result?}

We implemented the correlated-calls transformation and the IDE algorithm in Scala,
on top of the WALA framework for static analysis of JVM bytecode~\cite{fink2012wala}.
We also report on preliminary experiments in which our correlated-calls transformation
is applied to an IFDS formulation of a simple taint analysis. Our results show that
solving the resulting IDE problem avoids infeasible paths due to correlated calls as
expected.  
\mtodo{As discussed on the phone, we want to talk more about the results here.}

>>>>

Data-flow analysis computes an approximation of how values may flow  
through a program, and has applications in compiler optimization, 
programming tools, and computer security, and many other areas.
Data-flow analyses operate on \textit{control-flow graphs} (CFGs)%
\mtodo{If we introduce this abbreviation, should we replace the uses of
``control-flow graph'' with ``CFG'' in the paper?}
that model 
the order in which the instructions of a program are executed. This is
typically done by associating \textit{flow functions} that represent how 
data is propagated with the edges of the control-flow graph.  A 
\textit{confluence operator} specifies how the data facts that have been 
computed along different paths should be merged when the paths join.

@@

We present a formalization of the transformation and prove its correctness
as follows. First, we derive an ``Equivalence IDE problem''%
\mtodo{The IDE paper already states that each IFDS problem can be solved with IDE.
I wonder if this sounds like the equivalence transformation was our contribution.
We just use the equivalence transformation to explain the CC transformation, and for the
proofs.
}
from the original 
IFDS problem by associating the identity environment functions with all edges,
and show that the solution to this Equivalence IDE problem can be mapped back
to the solution of the original IFDS problem. Then, we derive a ``Correlated Calls 
IDE Problem'' from the original IFDS problem and show how a solution to this problem 
can be mapped to the solution to the original IFDS problem, but also how a
more precise IFDS result can be derived from it. 
\otodo{Do we need to mention the imprecise map from the CC-IDE result to the IFDS result? It
doesn't really seem relevant, and could be confusing.}
We also show that the correlated-calls analysis is sound, i.e., that it never considers concrete execution paths as infeasible.
This is illustrated schematically
in Figure~\ref{fig:transformations}%
\footnote{ 
  The labels $\result$, $\backEq$, $\backCC$, $\transEq$, and $\transCC$ on edges in 
  Figure~\ref{fig:transformations} reflect a number of mappings and projections 
  that will be defined in Section~\TODO{ref} and that can be ignored here.
}.  
\ftodo{would it make sense to add an edge in Figure~\ref{fig:transformations} from the IFDS problem to the IFDS result?}

We implemented the correlated-calls transformation and the IDE algorithm in Scala,
on top of the WALA framework for static analysis of JVM bytecode~\cite{fink2012wala}.
We also report on preliminary experiments in which our correlated-calls transformation
is applied to an IFDS formulation of a simple taint analysis. Our results show that
solving the resulting IDE problem avoids infeasible paths due to correlated calls as
expected.
